Kindly Bent to Free Us

### 2019-11-18 kindly/infer/exampleconstr.tex
Put the constraint example in the body, and push all the constrain theory out.

We now demonstrate the constraint solving
algorithm presented in \cref{infer:solving}
on the expression $\lam{f}{\lam{x}{(\app{f}{x},x)}}$.
The application of our inference algorithm yields the following constraints:
%
\begin{align*}
  \E &= (\tvar_f : \kvar_f)
  (\tvar_x : \kvar_x)\dots\\
  C &= (\tvar_f \leq \gamma \tarr{\kvar_1} \beta )
  \Cand
  (\gamma \leq \tvar_x)
  \Cand
  (\beta * \tvar_x \leq \alpha_r)
  \Cand
  (\kvar_x \leq \kun)
\end{align*}

@@

and the following kind constraints: 
\[\begin{aligned}
  \Cproj{\kvar_r}\Cproj{\kvar_f}\Cproj{\kvar_x}{}&
  (\kvar_x \leq \kun)
  \Cand
  (\kvar_\gamma \leq \kvar_x)
  \Cand
  (\kvar_x \leq \kvar_r)
  \Cand\\
  &(\kvar_\beta \leq \kvar_r)
  \Cand
  (\kvar_3 \leq \kvar_f)
  \Cand
  (\kvar_f \leq \kvar_1)
\end{aligned}\]
Since $\kvar_r$, $\kvar_f$ and $\kvar_x$ are not present in the type skeleton,
they are existentially quantified.

We translate these constraints into a relation whose graph
is shown in \cref{example:graph}.
%
The algorithm then proceeds as follow:
\begin{itemize}[noitemsep]
\item[1] From the constraints above, we deduce the graph shown
  with plain arrows.
\item[2] We add all the dashed arrows by saturating
  lattice inequalities. For clarity, we only show $\kun$.
\item[4] We identify the connected component circled in
  {\color{green} green}.
  We deduce $\kvar_\gamma = \kvar_x = \kun$
\item[5] We take the transitive closure, which adds the
  arrow in {\color{blue} blue} from $\kvar_1$ to $\kvar_3$.
\item[6] We remove the remaining existentially quantified node in {\color{red} red}, $\kvar_r$ and $\kvar_f$.
\item[7, 8, 9] We clean up the graph, and obtain the graph shown
  in \cref{example:graph:final}.
  We deduce $\kvar_3 \leq \kvar_1$.
\end{itemize}

@@

\begin{figure}[tbp]
  \centering
  \begin{tikzpicture}[every edge/.style=[link,->,>=latex,thick]]
    \node (U) {$\kun$} ;
    \node[above left=of U] (x) {$\kvar_x$} ;
    \node[above=of x] (r) {$\kvar_r$};
    \node[left=of x] (g) {$\kvar_\gamma$} ;
    \node[right=of x] (b) {$\kvar_\beta$} ;
    \node[right=of b] (3) {$\kvar_3$} ;
    \node[above=of 3] (f) {$\kvar_f$} ;
    \node[above=of f] (1) {$\kvar_1$} ;

    \begin{scope}[on background layer]
      \node[fill=green!20,draw=green,
      inner sep=-1mm,ellipse,rotate fit=-20,fit=(U) (x) (g)] {};
      \node[fill=red!20,draw=red, inner sep=-2pt, circle,fit=(r)] {};
      \node[fill=red!20,draw=red, inner sep=-2pt, circle,fit=(f)] {};
    \end{scope}
  \end{tikzpicture}
    \caption{Graph representing the example constraints}
    \label{example:graph}

  \begin{tikzpicture}[every edge/.style=[link,->,>=latex,thick]]
    \node (b) {$\kvar_\beta$} ;
    \node[right=of b] (3) {$\kvar_3$} ;
    \node[above=of 3] (f) {} ;
    \node[above=of f] (1) {$\kvar_1$} ;

  \end{tikzpicture}
    \caption{Final state of the graph}
    \label{example:graph:final}
\end{figure}

>>>>

\label{solving:example}

\begin{figure}[tbp]
  \centering
  \begin{tikzpicture}[node distance=7mm,every edge/.style=[link,->,>=latex,thick]]
    \node (U) {$\kun$} ;
    \node[above left=of U] (x) {$\kvar_x$} ;
    \node[above=of x] (r) {$\kvar_r$};
    \node[left=of x] (g) {$\kvar_\gamma$} ;
    \node[right=of x] (b) {$\kvar_\beta$} ;
    \node[right=of b] (3) {$\kvar_3$} ;
    \node[above=of 3] (f) {$\kvar_f$} ;
    \node[above=of f] (1) {$\kvar_1$} ;

    \node at (-0.5,-0.5) {Before};

    \begin{scope}[on background layer]
      \node[fill=green!20,draw=green,
      inner sep=-1pt,ellipse,rotate fit=-20,fit=(U.south) (x) (g)] {};
      \node[fill=red!20,draw=red, inner sep=-2pt, circle,fit=(r)] {};
      \node[fill=red!20,draw=red, inner sep=-2pt, circle,fit=(f)] {};
    \end{scope}
  \end{tikzpicture}~
  \vrule~
  \begin{tikzpicture}[node distance=7mm,every edge/.style=[link,->,>=latex,thick]]
    \node (b) {$\kvar_\beta$} ;
    \node[right=of b] (3) {$\kvar_3$} ;
    \node[above=of 3] (f) {} ;
    \node[above=of f] (1) {$\kvar_1$} ;

    \node at (0.5,-1.5) {After};
  \end{tikzpicture}
    \caption{Graph representing the example constraints}
    \label{example:graph}
    % \caption{Final state of the graph}
    \label{example:graph:final}
\end{figure}

We consider the expression $\lam{f}{\lam{x}{(\app{f}{x},x)}}$.
The application of our inference algorithm yields the following constraints:
%
\begin{align*}
  \E &= (\tvar_f : \kvar_f)
  (\tvar_x : \kvar_x)\dots\\
  C &= (\tvar_f \leq \gamma \tarr{\kvar_1} \beta )
  \Cand
  (\gamma \leq \tvar_x)
  \Cand
  (\beta * \tvar_x \leq \alpha_r)
  \Cand
  (\kvar_x \leq \kun)
\end{align*}

@@

and the following kind constraints: 
\[\begin{aligned}
    % \Cproj{\kvar_r}\Cproj{\kvar_f}\Cproj{\kvar_x}{}
    &(\kvar_x \leq \kun)
    \Cand
    (\kvar_\gamma \leq \kvar_x)
    \Cand
    (\kvar_x \leq \kvar_r)
    \Cand\\
    &(\kvar_\beta \leq \kvar_r)
    \Cand
    (\kvar_3 \leq \kvar_f)
    \Cand
    (\kvar_f \leq \kvar_1)
\end{aligned}\]
% Since $\kvar_r$, $\kvar_f$ and $\kvar_x$ are not present in the type skeleton,
% they are existentially quantified.

We translate these constraints into a relation whose graph
is shown in \cref{example:graph}.
%
The algorithm then proceeds as follow:
\begin{itemize}[noitemsep]
\item From the constraints above, we deduce the graph shown
  with plain arrows on the left
\item We add all the dashed arrows by saturating
  lattice inequalities. For clarity, we only show $\kun$.
\item We identify the connected component circled in
  {\color{green} green}.
  We deduce $\kvar_\gamma = \kvar_x = \kun$
\item We take the transitive closure, which adds the
  arrow in {\color{blue} blue} from $\kvar_1$ to $\kvar_3$.
\item We remove the remaining nodes not present in the type skeleton (colored in {\color{red} red}): $\kvar_r$ and $\kvar_f$.
\item We clean up the graph (transitive reduction, remove unneeded constants, \dots),
  and obtain the graph shown on the right.
  We deduce $\kvar_3 \leq \kvar_1$.
\end{itemize}

@@

### 2019-11-18 kindly/infer/infer.tex
Put the constraint example in the body, and push all the constrain theory out.

We define the constraint solver in terms of an arbitrary commutative bounded
lattice $(\mathcal L, \lk_\Lat)$, i.e.,
a lattice which has a minimal and a maximal element ($l^\top$ and $l^\bot$)
and where meet and joins are commutative.
We write lattice elements as $l$ and $\glb_i l_i$ (resp. $\lub_i l_i$)
for the greatest lower bound (resp. least upper bound) in $\mathcal L$.
The lattice for \lang (see \cref{sdtyping}) is a bounded lattice with
$l^\top = \klin_\infty$ and $l^\bot = \kun_0$.

Let $\CL$ be the set of constraints in such a lattice $\mathcal L$.
The full grammar of constraints is shown in \cref{grammar:constraint}.
Constraints are made of kind inequalities, conjunctions and
projections, as shown previously, along with type unification
constraints. Since types might contain kinds (for instance, on the arrows),
type unification is oriented and written as $\leq$.
For simplicity, we consider all type constructors
invariant in their parameters
and define $\Ceq{\tau}{\tau'}$ as $\Cleq{\tau}{\tau'} \wedge \Cleq{\tau'}{\tau}$.

As before, entailment is denoted by $\entail{C}{D}$,
where $D$ is a consequence of the constraints $C$.
We say that $C$ and $D$ are equivalent, $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
The base rules for entailment are given in \cref{rules:entail}.
The rules for classic Herbrand unification and
projection operators are omitted for brevity (see
\citep{DBLP:journals/tapos/OderskySW99} for
details).

\begin{figure*}[!bt]
  \centering
  \begin{minipage}{0.32\linewidth}
    \begin{align*}
      C &::=  C_1 \Cand C_2  \tag{Conjunction}\\
        &\mid \Cleq{\tau_1}{\tau_2} \mid \Cleq{k_1}{k_2} \tag{Inequalities} \\
        &\mid \Cproj{\tvar}{C}
        \mid \Cproj{\kvar}{C} \tag{Projections}
    \end{align*}
    \caption{The constraint language}
    \label{grammar:constraint}
  \end{minipage}\hfill
  \begin{minipage}{0.66\linewidth}
    \begin{mathpar}
      \inferrule
      {\forall i,\ \entail{C}{\Cleq{l_i}{k}}}
      {\entail{C}{\Cleq{\wedge_i l_i}{k}}}
      \and
      \inferrule
      {\forall i,\ \entail{C}{\Cleq{k}{l_i}}}
      {\entail{C}{\Cleq{k}{\vee_i l_i}}}
      \and
      \inferrule
      { \forall i,\ \entail{C}{\Ceq{\tau_i}{\tau_i}}\\
      }
      { \entail{C}{\Cleq{\tapp{\tcon}{(\tau_i)}}{\tapp{\tcon}{(\tau'_i)}}} }
      \and
      \inferrule{l \lk_\Lat l'}{\entail{}{\Cleq{l}{l'}}}
      \and
      \inferrule
      { \entail{C}{\Cleq{\tau'_1}{\tau_1}}\\
        \entail{C}{\Cleq{\tau_2}{\tau'_2}}\\
        \entail{C}{\Cleq{k}{k'}}
      }
      { \entail{C}{\Cleq{\tau_1\tarr{k}\tau_2}{\tau'_1\tarr{k'}\tau'_2}} }
    \end{mathpar}
    \caption{Base entailment rules -- $\entail{\inP{C}}{\inP{D}}$ }
    \label{rules:entail}
  \end{minipage}
\end{figure*}

We define the set of solved formed
$\mathcal S$ as the quotient set of $\CL$ by $\equivC$.
We will show later that such constraints are in fact only composed of
kind inequalities, and thus correspond to the syntactic constraints
used in type and kind schemes.
%
We now define our normalization procedure $\normalize{C_0}{\unif_0}$, where
$C_0\in \CL$ is a set of constraints and $\unif_0$ is a substitution.
It returns a constraint $C \in \mathcal S$ in
solved form and a unifier $\unif$.
The main idea of the algorithm is to first remove all the type equalities
by using regular Herbrand unification. After that, we only have
a set of inequalities among kinds, which we can consider as a relation.
We can then saturate the relation,
unify all kinds that are in the same equivalence classes to obtain
a most general unifier on kind variables,
remove all existentially quantified variables and
then minimize back the relation and apply various
simplification rules to make the resulting type easier to understand to users.

More precisely, we apply the following steps:
\begin{enumerate}
\item Solve all type equality constraints through Herbrand unification and
  gather all existential quantifications at the front of the constraint.
  We obtain a constraint $C^k = \exists \Multi\kvar,\ \Cleq{k_j}{k'_j}_j$ and
  a substitution $\unif_\tau$.

  We write $\mathcal R$ for the relation $\Cleq{k_j}{k'_j}_j$,
  $\mathcal G$ the underlying directed graph and $V$ its vertices.

\item Saturate the lattice equalities in $\mathcal R$.

  More precisely, for each kind variable $\kvar \in V$,
  for each constant $l_i$ (resp. $l_j$) such that
  there is a path from $l_i$ to $\kvar$ (resp. from $\kvar$ to $l_j$) in $\mathcal G$,
  add an edge from $\lub l_i$ to $\kvar$
  (resp. from $\kvar$ to $\glb l_j$).
  This step is well defined since $\mathcal L$ is a bounded lattice
  and $\lub\emptyset$ and $\glb\emptyset$ are well defined.

  We also complement $\mathcal R$ with $(\leq)$ by adding an edge
  between related constants.
\item
  At this point, we can easily check for satisfiability: A constraint
  is satisfiable (in the given environment) if and only if,
  for any constants $l_1$ and $l_2$ such that
  there is a path from $l_1$ to $l_2$ in $\mathcal G$, then $l_1\lk_\Lat l_2$.
  If this is not the case, we return \textbf{fail}.

\item For each strongly connected component in $\mathcal G$, unify all its vertices and replace it by a representative.
  We write $\unif_k$ for the substitution that replaces a kind variable by
  its representative.
  The representative of a strongly connected component $g$ can be determined as follows:
  \begin{itemize}
  \item If $g$ does not contain any constant, then the representative
    is a fresh kind variable.
  \item If $g$ contains exactly one constant, it is the representative.
  \item Otherwise, the initial constraint $C_0$ is not satisfiable.
  \end{itemize}
  Note that this step will also detect all unsatisfiable constraints.
\item Take the transitive closure of $\mathcal R$.
\item Remove all the vertices corresponding to the kind variables $\Multi\kvar$
  that are existentially quantified in $C^k$.
\item Take the transitive reduction of $\mathcal R$.
\item Remove the extremums of $\mathcal L$ and the edges of $(\leq)$
  from $\mathcal R$.
\item Return $C = \left\{ k \leq k' \mid k \operatorname{\mathcal R}k' \right\}$
  and $\unif =  \unif_\tau \meet \unif_k$.
\end{enumerate}

An example of this algorithm in action is shown in \cref{solving:example}.
Our algorithm is complete, computes principal normal forms
(cf \cref{appendix:constraints}),
and already simplifies constraints significantly
(thanks to steps 6, 7 and 8).
It can be extended with further simplification phases.
In particular, our implementation and all the signatures presented in
\cref{motivation} use a variance-based simplification
where all covariant (resp. contravariant) variables are replaced by their
lower (resp. upper) bounds.
All the simplification mechanisms presented
here, including the variance-based one, are complete.
It is also possible to add ``best-effort'' simplification
rules which help reduce the size of inferred signatures even further
\citep{DBLP:conf/aplas/Simonet03}.

>>>>

%
% \begin{figure}[tb]
%   \centering
%   \begin{align*}
%     C &::= \Cleq{\tau_1}{\tau_2}
%         \mid \Cleq{k_1}{k_2}
%         \mid C_1 \Cand C_2
%         \mid \Cproj{\tvar}{C}
%         \mid \Cproj{\kvar}{C}
%   \end{align*}
%   \caption{The constraint language}
%   \label{grammar:constraint}
%   % \input{infer/lattice}
%   % \caption{Lattice inequalities -- $k \lk_\Lat k'$}
%   \input{infer/entails}
%   \caption{Base entailment rules -- $\entail{\inP{C}}{\inP{D}}$ }
%   \label{rules:entail}
% \end{figure}

For concision, we demonstrate our constraint solving
algorithm on an example. The complete constraint system is
defined in \cref{appendix:constraints}.
%
\input{infer/exampleconstr}

Our algorithm respect the properties
of HM(X), computes principal normal forms
and simplifies constraints significantly.
All the simplification mechanisms presented
here, including the variance-based one, are complete.
It is also possible to add ``best-effort'' simplification
rules which help reduce the size of inferred signatures even further
\citep{DBLP:conf/aplas/Simonet03}.

### 2019-11-18 kindly/sdtyping/extra.tex
Put the constraint example in the body, and push all the constrain theory out.

\section{Constraints}
\label{appendix:constraints}
%
\begin{figure}[tb]
  \centering
  \begin{align*}
    C &::= \Cleq{\tau_1}{\tau_2}
        \mid \Cleq{k_1}{k_2}
        \mid C_1 \Cand C_2
        \mid \Cproj{\tvar}{C}
        \mid \Cproj{\kvar}{C}
  \end{align*}
  \caption{The constraint language}
  \label{grammar:constraint}
  % \input{infer/lattice}
  % \caption{Lattice inequalities -- $k \lk_\Lat k'$}
  \input{infer/entails}
  \caption{Base entailment rules -- $\entail{\inP{C}}{\inP{D}}$ }
  \label{rules:entail}
\end{figure}

@@

Our system is defined in \cref{infer:solving}.
As before, we write $\CL$ for our constraint system for an arbitrary
total bounded lattice $(\mathcal L, \lk_\Lat)$.
Entailment is written $\entail{\inP{C}}{\inP{D}}$,
where $D$ is a consequence of the
constraints $C$.
We say that $C$ and $D$ are equivalent, written as $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
Following HM(X), we define a cylindric constraint system as
a system which respects the following properties for any variable $x$:
\begin{align*}
  \entail{C&}{\Cproj{x}{C}}
  &\entail{C}{D} &\implies \entail{\Cproj{x}C}{\Cproj{x}D}\\
  \Cproj{x}{(C\wedge \Cproj{x}D)}&\equivC \Cproj{x}{C} \wedge \Cproj{x}D
  & \Cproj{x}{\Cproj{y}D}&\equivC \Cproj{y}{\Cproj{x}D}
\end{align*}

Furthermore, we define a term rewriting system as a system where, for every
types $\tau$,$\tau'$, there exists an equality predicates $\Ceq{\tau}{\tau'}$
which is symmetric, reflexive, transitive, stable under substitution and such that,
for any predicate $P$:
\begin{align*}
  &\entail{\Ceq{x}{y}\wedge \Cproj{x}C \wedge \Ceq{x}{y}}{C}\\
  &\subst{x}{\tau}{P} \equivC \Cproj{x}P \wedge \Ceq{x}{\tau}
                        \text{ where } x\notin\fv{\tau}
\end{align*}
$\CL$ is defined as the smallest cylindric term constraint system that
satisfies the axiom shown in \cref{rules:entail}.
We define  $\Ceq{\tau}{\tau'}$ as $\Cleq{\tau}{\tau'} \wedge \Cleq{\tau'}{\tau}$.
%
We write $\SC$ for the set of solved forms,
which can be used inside type and kind schemes.
We define $\SC$ as $\CL$ quotiented by the relation $\equivC$.
%
The normalization function is described in \cref{infer:solving}.

\subsection{Example of constraint solving and simplification}
\label{solving:example}

\input{infer/exampleconstr}

@@

Our goal is to prove that $\CL$ supports all the properties necessary for
principal type inference, as defined by HM(X).
We first prove that constraint solving
does compute normal forms, and that such normal forms are unique.

>>>>

\section{Constraints}
\label{appendix:constraints}

@@

We place our constraint system in a more general setting.
We define the constraint solver in terms of an arbitrary commutative bounded
lattice $(\mathcal L, \lk_\Lat)$, i.e.,
a lattice which has a minimal and a maximal element ($l^\top$ and $l^\bot$)
and where meet and joins are commutative.
We write lattice elements as $l$ and $\glb_i l_i$ (resp. $\lub_i l_i$)
for the greatest lower bound (resp. least upper bound) in $\mathcal L$.
The lattice for \lang (see \cref{sdtyping}) is a bounded lattice with
$l^\top = \klin_\infty$ and $l^\bot = \kun_0$.

% Our system is defined in \cref{infer:solving}.
% As before, we write $\CL$ for our constraint system for an arbitrary
% total bounded lattice $(\mathcal L, \lk_\Lat)$.
% Entailment is written $\entail{\inP{C}}{\inP{D}}$,
% where $D$ is a consequence of the
% constraints $C$.
% We say that $C$ and $D$ are equivalent, written as $C \equivC D$,
% when $\entail{C}{D}$ and $\entail{D}{C}$.

\begin{figure}[tb]
  \centering
  \begin{align*}
    C &::= \Cleq{\tau_1}{\tau_2}
        \mid \Cleq{k_1}{k_2}
        \mid C_1 \Cand C_2
        \mid \Cproj{\tvar}{C}
        \mid \Cproj{\kvar}{C}
  \end{align*}
  \caption{The constraint language}
  \label{grammar:constraint}
  % \input{infer/lattice}
  % \caption{Lattice inequalities -- $k \lk_\Lat k'$}
  \input{infer/entails}
  \caption{Base entailment rules -- $\entail{\inP{C}}{\inP{D}}$ }
  \label{rules:entail}
\end{figure}

Let $\CL$ be the set of constraints in such a lattice $\mathcal L$.
The full grammar of constraints is shown in \cref{grammar:constraint}.
Constraints are made of kind inequalities, conjunctions and
projections along with type unification
constraints. Since types might contain kinds (for instance, on the arrows),
type unification is oriented and written as $\leq$.
For simplicity, we consider all type constructors
invariant in their parameters
and define $\Ceq{\tau}{\tau'}$ as $\Cleq{\tau}{\tau'} \wedge \Cleq{\tau'}{\tau}$.

Entailment is denoted by $\entail{C}{D}$,
where $D$ is a consequence of the constraints $C$.
We say that $C$ and $D$ are equivalent, $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
% The base rules for entailment are given in \cref{rules:entail}.
% The rules for classic Herbrand unification and
% projection operators are omitted for brevity (see
% \citep{DBLP:journals/tapos/OderskySW99} for
% details).
%
Following HM(X), we define a cylindric constraint system as
a system which respects the following properties for any variable $x$:
\begin{align*}
  \entail{C&}{\Cproj{x}{C}}
  &\entail{C}{D} &\implies \entail{\Cproj{x}C}{\Cproj{x}D}\\
  \Cproj{x}{(C\wedge \Cproj{x}D)}&\equivC \Cproj{x}{C} \wedge \Cproj{x}D
  & \Cproj{x}{\Cproj{y}D}&\equivC \Cproj{y}{\Cproj{x}D}
\end{align*}

Furthermore, we define a term rewriting system as a system where, for every
types $\tau$,$\tau'$, there exists an equality predicates $\Ceq{\tau}{\tau'}$
which is symmetric, reflexive, transitive, stable under substitution and such that,
for any predicate $P$:
\begin{align*}
  &\entail{\Ceq{x}{y}\wedge \Cproj{x}C \wedge \Ceq{x}{y}}{C}\\
  &\subst{x}{\tau}{P} \equivC \Cproj{x}P \wedge \Ceq{x}{\tau}
                        \text{ where } x\notin\fv{\tau}
\end{align*}
$\CL$ is defined as the smallest cylindric term constraint system that
satisfies the axiom shown in \cref{rules:entail}.

% \begin{figure*}[!bt]
%   \centering
%   \begin{minipage}{0.32\linewidth}
%     \begin{align*}
%       C &::=  C_1 \Cand C_2  \tag{Conjunction}\\
%         &\mid \Cleq{\tau_1}{\tau_2} \mid \Cleq{k_1}{k_2} \tag{Inequalities} \\
%         &\mid \Cproj{\tvar}{C}
%         \mid \Cproj{\kvar}{C} \tag{Projections}
%     \end{align*}
%     \caption{The constraint language}
%     \label{grammar:constraint}
%   \end{minipage}\hfill
%   \begin{minipage}{0.66\linewidth}
%     \begin{mathpar}
%       \inferrule
%       {\forall i,\ \entail{C}{\Cleq{l_i}{k}}}
%       {\entail{C}{\Cleq{\wedge_i l_i}{k}}}
%       \and
%       \inferrule
%       {\forall i,\ \entail{C}{\Cleq{k}{l_i}}}
%       {\entail{C}{\Cleq{k}{\vee_i l_i}}}
%       \and
%       \inferrule
%       { \forall i,\ \entail{C}{\Ceq{\tau_i}{\tau_i}}\\
%       }
%       { \entail{C}{\Cleq{\tapp{\tcon}{(\tau_i)}}{\tapp{\tcon}{(\tau'_i)}}} }
%       \and
%       \inferrule{l \lk_\Lat l'}{\entail{}{\Cleq{l}{l'}}}
%       \and
%       \inferrule
%       { \entail{C}{\Cleq{\tau'_1}{\tau_1}}\\
%         \entail{C}{\Cleq{\tau_2}{\tau'_2}}\\
%         \entail{C}{\Cleq{k}{k'}}
%       }
%       { \entail{C}{\Cleq{\tau_1\tarr{k}\tau_2}{\tau'_1\tarr{k'}\tau'_2}} }
%     \end{mathpar}
%     \caption{Base entailment rules -- $\entail{\inP{C}}{\inP{D}}$ }
%     \label{rules:entail}
%   \end{minipage}
% \end{figure*}

We define the set of solved formed
$\mathcal S$ as the quotient set of $\CL$ by $\equivC$.
We will show later that such constraints are in fact only composed of
kind inequalities, and thus correspond to the syntactic constraints
used in type and kind schemes.
%
We now define our normalization procedure $\normalize{C_0}{\unif_0}$, where
$C_0\in \CL$ is a set of constraints and $\unif_0$ is a substitution.
It returns a constraint $C \in \mathcal S$ in
solved form and a unifier $\unif$.
The main idea of the algorithm is to first remove all the type equalities
by using regular Herbrand unification. After that, we only have
a set of inequalities among kinds, which we can consider as a relation.
We can then saturate the relation,
unify all kinds that are in the same equivalence classes to obtain
a most general unifier on kind variables,
remove all existentially quantified variables and
then minimize back the relation and apply various
simplification rules to make the resulting type easier to understand to users.

More precisely, we apply the following steps:
\begin{enumerate}
\item Solve all type equality constraints through Herbrand unification and
  gather all existential quantifications at the front of the constraint.
  We obtain a constraint $C^k = \exists \Multi\kvar,\ \Cleq{k_j}{k'_j}_j$ and
  a substitution $\unif_\tau$.

  We write $\mathcal R$ for the relation $\Cleq{k_j}{k'_j}_j$,
  $\mathcal G$ the underlying directed graph and $V$ its vertices.

\item Saturate the lattice equalities in $\mathcal R$.

  More precisely, for each kind variable $\kvar \in V$,
  for each constant $l_i$ (resp. $l_j$) such that
  there is a path from $l_i$ to $\kvar$ (resp. from $\kvar$ to $l_j$) in $\mathcal G$,
  add an edge from $\lub l_i$ to $\kvar$
  (resp. from $\kvar$ to $\glb l_j$).
  This step is well defined since $\mathcal L$ is a bounded lattice
  and $\lub\emptyset$ and $\glb\emptyset$ are well defined.

  We also complement $\mathcal R$ with $(\leq)$ by adding an edge
  between related constants.
\item
  At this point, we can easily check for satisfiability: A constraint
  is satisfiable (in the given environment) if and only if,
  for any constants $l_1$ and $l_2$ such that
  there is a path from $l_1$ to $l_2$ in $\mathcal G$, then $l_1\lk_\Lat l_2$.
  If this is not the case, we return \textbf{fail}.

\item For each strongly connected component in $\mathcal G$, unify all its vertices and replace it by a representative.
  We write $\unif_k$ for the substitution that replaces a kind variable by
  its representative.
  The representative of a strongly connected component $g$ can be determined as follows:
  \begin{itemize}
  \item If $g$ does not contain any constant, then the representative
    is a fresh kind variable.
  \item If $g$ contains exactly one constant, it is the representative.
  \item Otherwise, the initial constraint $C_0$ is not satisfiable.
  \end{itemize}
  Note that this step will also detect all unsatisfiable constraints.
\item Take the transitive closure of $\mathcal R$.
\item Remove all the vertices corresponding to the kind variables $\Multi\kvar$
  that are existentially quantified in $C^k$.
\item Take the transitive reduction of $\mathcal R$.
\item Remove the extremums of $\mathcal L$ and the edges of $(\leq)$
  from $\mathcal R$.
\item Return $C = \left\{ k \leq k' \mid k \operatorname{\mathcal R}k' \right\}$
  and $\unif =  \unif_\tau \meet \unif_k$.
\end{enumerate}

An example of this algorithm in action is shown in \cref{solving:example}.
Our algorithm is complete, computes principal normal forms,
and already simplifies constraints significantly
(thanks to steps 6, 7 and 8).
It can be extended with further simplification phases.
In particular, our implementation and all the signatures presented in
\cref{motivation} use a variance-based simplification
where all covariant (resp. contravariant) variables are replaced by their
lower (resp. upper) bounds.
All the simplification mechanisms presented
here, including the variance-based one, are complete.
It is also possible to add ``best-effort'' simplification
rules which help reduce the size of inferred signatures even further
\citep{DBLP:conf/aplas/Simonet03}.

% \subsection{Example of constraint solving and simplification}
% \label{solving:example}

% \input{infer/exampleconstr}

@@

We now prove that $\CL$ supports all the properties necessary for
principal type inference, as defined by HM(X).
We first prove that constraint solving
does compute normal forms, and that such normal forms are unique.
